A component-based DSL for multi-stencil programs

### 2015-07-22 WOLFHPC15/article/concept.tex
wolfhpc15 22_07

%----------------------------------------
\subsection{Multi-stencil programs}
\label{sect:multistencil}
%----------------------------------------
To numerically solve a set of PDEs, iterative methods (finite difference, finite volume or finite element methods) are frequently used to approximate the solution through a discretized (step by step) phenomena. Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied onto a mesh (space discretization). In other words, in a mesh-based numerical simulation, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain (the mesh). Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}~\cite{spaaTangCKLL11}.

If $x$ represents an element of the mesh, if $N(x)$ is a neighborhood shape around the element $x$, and if $u_t(x)$ is the data (or quantity) to write. Moreover if $V_t(x)$ is the set of data to simulate such that $V_t(x) \cap u_t(x) = \emptyset$, a stencil computation is of the following form

\begin{equation} 
\{u_{t-1}(x), u_{t-1}(N(x)), V_t(x), V_t(N(x)\} \longmapsto u_{t}(x).
\label{space}
\end{equation}

@@

A real case numerical simulation is most of the time composed of a set of numerical schemes (stencils computations), with one or more stencil shape (neighboorhood shape), and onto one or more data. Moreover, a numerical simulation also performs additionnal auxiliary computations which do not involve neighborhood, called local computations. In this paper we target this kind of complex simulations with more than one stencil computation and also additionnal auxiliary computations, that we call \emph{multi-stencil programs}. The work presented produces a general parallelization structure of the overall simulation, while most stencil solutions produce optimized codes for a single stencil code~\cite{citeulike12258902, spaaTangCKLL11, Giles2011}. As it will be detailed in the related work, this work is complementary to most stencil compilers and implicit parallelism languages.

\textbf{Do I need the following definitions?}
\begin{mydef}
A \textit{stencil computation} is defined as a quadruplet $s(R,w,\text{exp},\mathcal{N})$,
\end{mydef}
where $R$ is the set of data read by the computation, $w$ is the data written by the computation, \textit{exp} is the numerical expression to compute $w$, and $\mathcal{N}$ is a neighborhood shape of the stencil computation.

@@

>>>>

%----------------------------------------
\subsection{Stencil codes}
\label{sect:stencil}
%----------------------------------------
To numerically solve a set of PDEs, iterative methods (finite difference, finite volume or finite element methods) are frequently used to approximate the solution through a discretized (step by step) phenomena. Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied onto a mesh (space discretization). In other words, in a mesh-based numerical simulation, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain (the mesh). Among those numerical computations is found a set of numerical schemes, also called \textit{stencil computations}~\cite{spaaTangCKLL11}.

\begin{figure}[!h]\begin{center}
  \resizebox{9cm}{!}{\includegraphics{./images/stencil.pdf}}
  \caption{Example of a stencil computation.}
  \label{fig:ex}
\end{center}\end{figure}

@@

%----------------------------------------
\subsection{Multi-stencil programs}
\label{sect:multistencil}
%----------------------------------------
A real case numerical simulation is most of the time composed of a set of numerical schemes (stencils computations), with one or more stencil shape (neighboorhood shape), and onto one or more data. Moreover, a numerical simulation also performs additionnal auxiliary computations which do not involve neighborhood, called local computations. In this paper we target this kind of complex simulations with more than one stencil computation and also additionnal auxiliary computations, that we call \emph{multi-stencil programs}. The work presented produces a general parallelization structure of the overall simulation, while most stencil solutions produce optimized codes for a single stencil code~\cite{citeulike12258902, spaaTangCKLL11, Giles2011}. As it will be detailed in the related work, this work is complementary to most stencil compilers and implicit parallelism languages.

\begin{mydef}
A \textit{stencil computation} is defined as a quadruplet $s(R,w,\text{exp},\mathcal{N})$,
\end{mydef}
where $R$ is the set of data read by the computation, $w$ is the data written by the computation, \textit{exp} is the numerical expression to compute $w$, and $\mathcal{N}$ is a neighborhood shape of the stencil computation.

@@

A multi-stencil program in this work is defined as a sextuplet
\begin{equation*} 
\mathcal{MSP}(T,\mathcal{M},\mathcal{D}_m,\mathcal{D}_c,\Delta,\Gamma)
\end{equation*}
